---
title: "HW6"
author: "Devan Becker"
date: "3/7/2019"
output: 
  html_document: 
    keep_md: yes
---
```{r}
library(tidyverse)
```

```{r}
library(skimr)
```

```{r}
library("RColorBrewer")
```

```{r}
library(gapminder)
```

```{r}
gapminder <-
  gapminder::gapminder
```

1. Explore the data using the various function you have learned. Is it tidy, are there any NA’s, what are its dimensions, what are the column names, etc.
```{r}
gapminder
```
##Data is Tidy.
```{r}
glimpse(gapminder)
```

```{r}
gapminder %>% 
  skimr::skim()
```

```{r}
gapminder %>% 
  summarize(number_nas= sum(is.na(gapminder)))
```
##No NA's.
```{r}
colnames(gapminder)
```

2. We are interested in the relationship between per capita GDP and life expectancy; i.e. does having more money help you live longer on average. Make a quick plot below to visualize this relationship.
```{r}
gapminder %>% 
  ggplot(aes(lifeExp, gdpPercap)) +
  geom_point() +
  scale_y_log10()
```
##More money equals longer life expectancy.
3. There is extreme disparity in per capita GDP. Rescale the x axis to make this easier to interpret. How would you characterize the relationship?
```{r}
gapminder %>% 
  ggplot(aes(lifeExp, gdpPercap)) +
  geom_point() +
  scale_y_log10() +
  geom_smooth(method = lm, se = TRUE)
```
##There is a strong positive relationship between GDP per capita and life expextancy.

4. This should look pretty dense to you with significant overplotting. Try using a faceting approach to break this relationship down by year.
```{r}
gapminder %>% 
  ggplot(aes(lifeExp, gdpPercap)) +
  geom_point(size = 0.3) +
  scale_y_log10() +
  geom_smooth(method = lm, se = TRUE) +
  facet_grid(~year)
```


5. Simplify the comparison by comparing only 1952 and 2007. Can you come to any conclusions?
```{r}
gapminder %>% 
  filter(year==1952 | year==2007) %>% 
  ggplot(aes(lifeExp, gdpPercap)) +
  geom_point(size = 0.8) +
  scale_y_log10() +
  geom_smooth(method = lm, se = TRUE) +
  facet_grid(~year)
```

6. Let’s stick with the 1952 and 2007 comparison but make some aesthetic adjustments. First try to color by continent and adjust the size of the points by population. Add + scale_size(range = c(0.1, 10), guide = "none") as a layer to clean things up a bit.
```{r}
gapminder %>% 
  filter(year==1952 | year==2007) %>% 
  ggplot(aes(lifeExp, gdpPercap, color = continent, size = pop)) +
  geom_point() +
  scale_y_log10() +
  facet_grid(~year) +
  scale_size(range = c(0.1, 10), guide = "none")
```

7. Although we did not introduce them in lab, ggplot has a number of built-in themes that make things easier. I like the light theme for these data, but you can see lots of options. Apply one of these to your plot above.
```{r}
library("RColorBrewer")
```

```{r}
gapminder %>% 
  filter(year==1952 | year==2007) %>% 
  ggplot(aes(lifeExp, gdpPercap, color = continent, size = pop)) +
  geom_point() +
  scale_y_log10() +
  facet_grid(~year) +
  scale_size(range = c(0.1, 10), guide = "none") +
  scale_colour_brewer(palette = "BrBG")
```


